\LARGE \bf On the Commissioning of the 12 GeV HMS Drift Chambers, Electronics/Computer Dead Time Monitoring and Overview of the D(e,e'p)n Experimental Run Plan

### 2017-10-29 NRF_Final_Report/report.tex
finished NRC 10-page report

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
\usepackage{url}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{graphicx}
\usepackage[scientific-notation=true]{siunitx}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage[top=0.6 in, bottom=0.7 in, left=0.7in, right=0.7in]{geometry}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{vector}
\usepackage{amsmath}
\usepackage{empheq}
\usepackage{amssymb}
\usepackage{tgtermes}
\usepackage{gensymb}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{epstopdf}
\usepackage[singlelinecheck=false]{caption}
\usepackage{datetime}
\usepackage[hyperfootnotes=false, pdfnewwindow, colorlinks=true]{hyperref}

@@

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins

\title{\LARGE \bf
  On the Commissioning of the 12 GeV HMS Drift Chambers, Electronics/Computer Dead Time Monitoring
  and Overview of the D(e,e'p)n Experimental Run Plan}

\author{Carlos Yero \\
October 30, 2017}

@@

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
      Three separate topics, all of equal importance, are briefly discussed. The new (12 GeV Era) HMS Drift Chambers are ready to be put in the HMS
      detector stack, in place of the old HMS Chambers. Several efficcinecy tests were performed on one of the chambers during the second week of October
      2017. The efficciecies were determined to be better than 99$\%$. The second chamber has not been tested yet, but it is
      expected to behave the same since both chambers were tested under similar conditions in the past. Dead time studies are currently in progress to
      determine how many physics events (triggers) are actually lost due to computer and electronic deadtime inherent in our experimental equipment. 
      There had been some technical issues found related to the computer livetime that are being addressed by the Jefferson Lab DAQ group. The experimental run
      plan of my thesis experiment, the electro-disintegration of Deuteron, is briefly discussed as the kinematics have slightly changed and new
      simulations had to be done. \\      
    \end{abstract}
  \end{@twocolumnfalse}
  ]

@@

On March 7-10 of 2017, a 5 $\mu$A electron beam was delivered to a BeO and Carbon targets for the first time to experimental

@@

Chambers. The chambers were constructed at Hampton University by Dr. Liguang Tang and his graduate students in 2016.

@@

at lower cost compared to the gas mixture that the chambers run on during an experiment, which is why it is preferred during the testing phase of the detector.}

@@

due to a bad connection between the sense wires and the discriminator cards. The second chamber is now ready to be transported to Hall C for further tests with the

@@

the number of physics events (triggers) are counted via nuclear electronic modules. These triggers ultimately get processed by the DAQ before being

@@

\subsection{Design and Operation of the Chambers}
The new HMS Drift Chambers were designed to be geometrically the same as the SHMS Drift Chambers. Each chamber consists of 6 wire planes and each wire
plane is located between two cathode planes. The wire planes consist of alternating field and sense wires. The U, U', V and V' planes consist of
96 sense wires each and are oriented 60$^{o}$ relative to the +y-coordinate. The X and X' planes consist of 102 sense wires and are oriented perpendicular to the x-axis (See
Figures \ref{fig:dc1_planes} and \ref{fig:dc_wires}).
\begin{figure}[h!]
  \centering
  \includegraphics[width=3.0in, height=2.5in]{dc2_tests/HMS_DC1_Planes.pdf}
  \caption{HMS Drift Chamber 1 wire planes.}
  \label{fig:dc1_planes}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=3.0in, height=1.5in]{dc2_tests/HMS_DC_Wires.pdf}
  \caption{HMS Drift Chamber wire orientation for planes a) U, U', b) V, V' and c) X, X' where
  the direction of the beam (+z) is into the plane of the paper.}
  \label{fig:dc_wires}
\end{figure}
The cathode planes and field wires are held at a negative high voltage and the sense wires are grounded which establihses a potential gradient causing an electric field
between high voltage and grounded wires. Each chamber is filled with a gas mixture\footnote{Argon is mixed with CO$_{2}$ or Ethane, where Argon is the ionizing agent the
particles interact with, and Ethane/CO$_{2}$ are the quenching elements to control avalanche produced by secondary ionization} of Argon with either CO$_{2}$ or Ethane.
As the particles traverse the chamber, the free electrons from ionized Argon drift towards the sense wires producing a detectable signal which is read out via discriminator
cards and into electronic modules.

@@

To determine the operatility of the HMS Drift Chambers, several cosmic tests were performed on the chambers in the Experimental Storage Building (ESB) before being transported

@@

  \includegraphics[width=4.0in, height=2.4in]{dc2_tests/gas_mix_current_drawn.pdf}

@@

bottom PMTs is defined as the \textit{trigger}. The efficiency of the \textit{i$^{th}$} plane, $\epsilon_{i}$, can then be defined as follows,
\epsilon_{i} = \frac{\# \text{ triggers that \textit{did} hit the \textit{i$^{th}$} plane } }{\# \text{ triggers that \textit{should} have hit the \textit{i$^{th}$} plane} } 

@@

  \includegraphics[width=3.8in, height=2.2in]{dc2_tests/hdc2_xmap.pdf}

@@

  \includegraphics[width=4.1in, height=2.2in]{dc2_tests/dc2_hvscan_thrs45.pdf}

@@

  \includegraphics[width=4.1in, height=2.2in]{dc2_tests/dc2_thrsscan_HV_1940V.pdf}

@@

\section{MATH}

Before you begin to format your paper, first write and save the content as a separate text file. Keep your text and graphic files separate until after the text has been formatted and styled. Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph. Do not add any kind of pagination anywhere in the paper. Do not number text heads-the template will do that for you.

Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:

\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract. Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined. Do not use abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}

@@

\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as Ò3.5-inch disk driveÓ.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ÒWb/m2Ó or Òwebers per square meterÓ, not Òwebers/m2Ó.  Spell out units when they appear in text: Ò. . . a few henriesÓ, not Ò. . . a few HÓ.
\item Use a zero before decimal points: Ò0.25Ó, not Ò.25Ó. Use Òcm3Ó, not ÒccÓ. (bullet list)

@@

\subsection{Equations}

The equations are an exception to the prescribed specifications of this template. You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font). To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled. Number equations consecutively. Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop. To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents. Italicize Roman symbols for quantities and variables, but not Greek symbols. Use a long dash rather than a hyphen for a minus sign. Punctuate equations with commas or periods when they are part of a sentence, as in

$$
\alpha + \beta = \chi \eqno{(1)}
$$

Note that the equation is centered using a center tab stop. Be sure that the symbols in your equation have been defined before or immediately following the equation. Use Ò(1)Ó, not ÒEq. (1)Ó or Òequation (1)Ó, except at the beginning of a sentence: ÒEquation (1) is . . .Ó

\subsection{Some Common Mistakes}
\begin{itemize}

\item The word ÒdataÓ is plural, not singular. \cite{Leo}
\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ÒoÓ.
\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ÒinsetÓ, not an ÒinsertÓ. The word alternatively is preferred to the word ÒalternatelyÓ (unless you really mean something that alternates).
\item Do not use the word ÒessentiallyÓ to mean ÒapproximatelyÓ or ÒeffectivelyÓ.
\item In your paper title, if the words Òthat usesÓ can accurately replace the word ÒusingÓ, capitalize the ÒuÓ; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ÒaffectÓ and ÒeffectÓ, ÒcomplementÓ and ÒcomplimentÓ, ÒdiscreetÓ and ÒdiscreteÓ, ÒprincipalÓ and ÒprincipleÓ.
\item Do not confuse ÒimplyÓ and ÒinferÓ.
\item The prefix ÒnonÓ is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ÒetÓ in the Latin abbreviation Òet al.Ó.
\item The abbreviation Òi.e.Ó means Òthat isÓ, and the abbreviation Òe.g.Ó means Òfor exampleÓ.

\section{USING THE TEMPLATE}

Use this sample document as your LaTeX source file to create your document. Save this file as {\bf root.tex}. You have to make sure to use the cls file that came with this distribution. If you use a different style file, you cannot expect to get required margins. Note also that when you are creating your out PDF file, the source file is only part of the equation. {\it Your \TeX\ $\rightarrow$ PDF filter determines the output file size. Even if you make all the specifications to output a letter file in the source - if you filter is set to produce A4, you will only get A4 output. }

It is impossible to account for all possible situation, one would encounter using \TeX. If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.

\subsection{Headings, etc}

Text heads organize the topics on a relational, hierarchical basis. For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic. If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced. Styles named ÒHeading 1Ó, ÒHeading 2Ó, ÒHeading 3Ó, and ÒHeading 4Ó are prescribed.

\subsection{Figures and Tables}

Positioning Figures and Tables: Place figures and tables at the top and bottom of columns. Avoid placing them in the middle of columns. Large figures and tables may span across both columns. Figure captions should be below the figures; table heads should appear above the tables. Insert figures and tables after they are cited in the text. Use the abbreviation ÒFig. 1Ó, even at the beginning of a sentence.

\begin{table}[h]
\caption{An Example of a Table}
\label{table_example}
\begin{center}
\begin{tabular}{|c||c|}
\hline
One & Two\\
\hline
Three & Four\\
\hline
\end{tabular}
\end{center}
\end{table}

   \begin{figure}[thpb]
      \centering
      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
}}
      %\includegraphics[scale=1.0]{figurefile}
      \caption{Inductance of oscillation winding on amorphous
       magnetic core versus DC bias magnetic field}
      \label{figurelabel}
   \end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader. As an example, write the quantity ÒMagnetizationÓ, or ÒMagnetization, MÓ, not just ÒMÓ. If including units in the label, present them within parentheses. Do not label axes only with units. In the example, write ÒMagnetization (A/m)Ó or ÒMagnetization {A[m(1)]}Ó, not just ÒA/mÓ. Do not label axes with a ratio of quantities and units. For example, write ÒTemperature (K)Ó, not ÒTemperature/K.Ó

\section{CONCLUSIONS}

A conclusion section is not required. Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion. A conclusion might elaborate on the importance of the work or suggest applications and extensions.

>>>>

\documentclass[letterpaper, 12 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
\usepackage{url}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{graphicx}
\usepackage[scientific-notation=true]{siunitx}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage[top=0.6 in, bottom=0.7 in, left=0.7in, right=0.7in]{geometry}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{vector}
\usepackage{amsmath}
\usepackage{empheq}
\usepackage{amssymb}
\usepackage{tgtermes}
\usepackage{gensymb}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{epstopdf}
\usepackage[singlelinecheck=false]{caption}
\usepackage{datetime}
\usepackage[hyperfootnotes=false, pdfnewwindow, colorlinks=true]{hyperref}

@@

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
\pagestyle{plain}
\title{\LARGE \bf
  On the Commissioning of the 12 GeV HMS Drift Chambers, Electronics/Computer Dead Time Monitoring
  and Overview of the D(e,e'p)n Experimental Run Plan}

\author{\normalsize Carlos Yero \\
  Major Advisor: Dr. Werner Boeglin\\
  October 30, 2017}

@@

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle

    \pagestyle{plain}
    \thispagestyle{plain}

    \begin{abstract}
      Three separate topics, all of equal importance, are briefly discussed. The new (12 GeV Era) HMS Drift Chambers are ready to be put in the HMS
      detector stack, in place of the old HMS Chambers. Several efficcinecy tests were performed on one of the chambers during the second week of October
      2017. The efficciecies were determined to be better than 99$\%$. The second chamber has not been tested yet, but it is
      expected to behave the same since both chambers were tested under similar conditions in the past. Dead time studies are currently in progress to
      determine how many physics events (triggers) are actually lost due to computer and electronic deadtime inherent in our experimental equipment. 
      There had been some technical issues found related to the computer livetime that are being addressed by the Jefferson Lab DAQ group. The experimental run
      plan of my thesis experiment, the electro-disintegration of Deuteron, is briefly discussed as the kinematics have slightly changed and new
      simulations had to be done. \\      
    \end{abstract}
  \end{@twocolumnfalse}
  ]

@@

\noindent On March 7-10 of 2017, a 5 $\mu$A electron beam was delivered for the first time to experimental

@@

Chambers. The chambers were constructed at Hampton University by Dr. Liguang Tang and his graduate students in 2016\cite{bishnu}.

@@

safer to use outside the Hall, compared than the gas mixture that the chambers run on during an experiment, which is why it is preferred during the testing phase of the detector.}

@@

due to a bad connection between the sense wires and external connector. The second chamber is now ready to be transported to Hall C for further tests with the

@@

the number of physics events (or triggers\footnote{A condition that is imposed on the detected signals to determine whether or not it corresponds to a physics event.}  are counted via nuclear electronic modules. These triggers ultimately get processed by the DAQ before being

@@

\subsection{Design and Operation of the Chambers}
\noindent The new HMS Drift Chambers were designed to be geometrically the same as the SHMS Drift Chambers. Each chamber consists of 6 wire planes and each wire
plane is located between two cathode planes. The wire planes consist of alternating field and sense wires. The U, U', V and V' planes consist of
96 sense wires each and are oriented 60$^{o}$ relative to the +y-coordinate. The X and X' planes consist of 102 sense wires and are oriented perpendicular to the x-axis (See
Figures \ref{fig:dc1_planes} and \ref{fig:dc_wires}).
\begin{figure}[h!]
  \centering
  \includegraphics[width=3.0in, height=2.5in]{dc2_tests/HMS_DC1_Planes.pdf}
  \caption{HMS Drift Chamber 1 wire planes.}
  \label{fig:dc1_planes}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=3.0in, height=1.5in]{dc2_tests/HMS_DC_Wires.pdf}
  \caption{HMS Drift Chamber wire orientation for planes a) U, U', b) V, V' and c) X, X' where
  the direction of the beam (+z) is into the plane of the paper.}
  \label{fig:dc_wires}
\end{figure}
The cathode planes and field wires are held at a negative high voltage and the sense wires are grounded which establihses a potential gradient causing an electric field
between high voltage and grounded wires. Each chamber is filled with a gas mixture\footnote{Argon is mixed with CO$_{2}$ or Ethane, where Argon is the ionizing agent the
particles interact with, and Ethane/CO$_{2}$ are the quenching elements to control avalanche produced by secondary ionization} of Argon with either CO$_{2}$ or Ethane.
As the particles traverse the chamber, the free electrons from ionized Argon drift towards the sense wires producing a detectable signal which is read out via discriminator
cards and into electronic modules.

@@

\noindent To determine the operatility of the HMS Drift Chambers, several cosmic tests were performed on the chambers in the Experimental Storage Building (ESB) before being transported

@@

  \includegraphics[width=3.5in, height=2.4in]{dc2_tests/gas_mix_current_drawn.pdf}

@@

bottom PMTs is defined as the \textit{trigger}. The efficiency of the \textit{i$^{th}$} plane, $\epsilon_{i}$, can then be defined as follows, \\
\begin{align*}
\epsilon_{i} = \frac{\# \text{ triggers that hit the \textit{i$^{th}$} plane } }{\# \text{ triggers that \textit{should} have hit the \textit{i$^{th}$} plane} } 
\end{align*} \\

@@

  \includegraphics[width=3.7in, height=2.2in]{dc2_tests/hdc2_xmap.pdf}

@@

  \includegraphics[width=3.8in, height=2.2in]{dc2_tests/dc2_hvscan_thrs45.pdf}

@@

  \includegraphics[width=3.8in, height=2.2in]{dc2_tests/dc2_thrsscan_HV_1940V.pdf}

@@

\section{Electronic and Computer Deadtime Studies}
\subsection{Dead Time Measurements Procedure}
\noindent The procedure to determine the computer and electronic dead times is currently in the development stage in Hall C. The general idea is to use a pulse generator to produce
a fixed frequency pulse and feed it as early as possible into the electronic modules being used in the experimental setup. This pulse will interfere wth the actual signals
produced by physics, but its rate can be set orders of magnitude smaller such that its interference will be minimal and will have insignificant contribution to the deadtime.
By counting the number of pulses generated before sending them to the electronics chain, compared to the number of pulses that actually make it to the end of the chain and
are accepted by the TI (Trigger Interface) module\cite{TI}, one may determine the dead time as follows:
\begin{equation} \label{deadtime}
  \text{total D.T.} = 1 - \underbrace{\frac{\# \text{ of accepted pulses}}{\# \text{ of generated pulses}}}_{\text{\normalsize total L.T.}} 
This is the total deadtime from the electronics and computer (TI module) that arises due to the finite processing time of the modules. From Eq. \ref{deadtime},  the total live time has contributions from the computer and electronics, which can
be calculated separately. \\
\indent The computer live time is calculated by the TI module since it has an
internal scaler which counts the number of input and accepted triggers, and takes the ratio. This calculation can also be done externally as a cross-check by counting the number of signals that make it to the
TI via an external scaler module, and comparing it to the number of accepted triggers which is outputted by the TI and determined during the data analysis stage. \\
\indent The electronics livetime cannot be determined directly,  but it can be determined by keeping the computer live time fixed near 100$\%$ and
increasing the pulse generator rate to probe the capability of the electronics modules to handle high rates. This way, if there is a decrease in the total live time which can be measured directly, then it must have been caused by the
electronics, since the computer live time is kept fixed. \\
\indent One might argue that such high rates can reduce the computer live time itself which is true, however, such high rates can be pre-scaled by the TI by orders
of magnitude, such that the computer live time is kept fixed. For example, the TI can be pre-scaled by 10$^{2}$ such that for a rate of 100,000 counts/sec. (100 KHz), the TI accepts 1 kHz.  Using this technique, one can investigate the
effects of high rates on the electronic modules without having a significant contribution from the computer live time. \\
\subsection{Basic Concepts on Computer Live Time}
\noindent To better understand how the computer live time is determined, I will introduce some basic concepts about how the TI module does this calculation. The main functionality of the TI is to process and distribute triggers and initiate data readout in fast electronics modules. Due to the finite processing time of these modules, the TI has a \textit{deadtime} in which it is unable to process any more triggers until after all other fast electronics modules have finished processing the current trigger. During this time, the TI generates a BUSY signal typically on the order of a few microseconds ($\mu$s) during which
all other input triggers are rejected. \\
\indent To understand computer live time experimentally, one can use a pulse generator to feed a controlled pulse into the TI and explore the effect of
various pulse rates on the deadtime. Based on the periodicity of the input pulse and the output BUSY signal width of the TI, three limiting cases arise:
  \includegraphics[width=3.4in, height=2.8in]{edtm/edtm_cases.pdf}
  \caption{Three limiting cases on TI computer live time based on the input signal.}
  \label{fig:edtm_cases}

@@

\item Case 1: The pulse periodicity falls outside the TI BUSY window so the ratio of accepted signals to input signals is unity and the live time is 100$\%$\\
\item Case 2: The pulse periodicity is equal to the width of the BUSY window, so every two input pulses, the first one wll open a BUSY window and the second will be blocked by the BUSY so the TI accepts only 1/2 the input signals and the live time is 50$\%$. \\
\item Case 3: The pulse periodicity is smaller than the BUSY window so the live time is 50$\%$ or less. 

@@

For the second and third case, a general expression for the computer live time can be obtained by expressing the pulse periodicity $T$ as a fraction
of the BUSY width, $w$. For a pulse period
\begin{equation}\label{eq:period}
T = \frac{1}{n}w, \text{ where n = 1, 2, 3, . . .} 
the computer live time can be shown to be 
\begin{equation}\label{eq:L.T.}
 \text{computer L.T. } = \frac{1}{n+1}
For example, if the period is one-half of the BUSY width, then the first pulse that comes initiates the BUSY, the second pulse will arrive half way, and
the third pulse will be at the end of the BUSY, so only 1/3 (33$\%$ live time) of the pulses will be accepted by the TI as the other two will be rejected since they did
not arrived outside the window. \\
\indent In the previous discussions, I have used the example of a pulse generator that produces periodic pulses in time. In reality, during a physics experiment, the physics events
that are detected follow poisson statistics, which is a random process. This means that pulses arrive random in time, and so it becomes impossible to predict exactly what the computer
live time will be (See Figure \ref{fig:poisson_pulse}).
  \includegraphics[width=3.1in, height=0.9in]{edtm/poisson_pulse.pdf}
  \caption{Random pulses in time due to physics events follow poisson statistics.}
  \label{fig:poisson_pulse}
\end{figure}\\
In Figure \ref{fig:poisson_pulse}, the first pulse initiates a BUSY signal, the following pulses then arrive at random times which may or may not be within the BUSY width so achieving a live
time of 100$\%$ becomes highly improbable during an experimental run.
\subsection{Computer Live Time Measurements}
\noindent Initial measurements of the computer live time were performed and the results of these measurements uncovered several issues associated with the
TI module and the DAQ which required the help from experts. The first set of measurements were done using a pulse generator at various rates up to 10 kHz. The pulses were fed to the electronics modules in both the HMS and SHMS and the BUSY signal widths were measured. The results are shown in Table \ref{table:cp_LT}. \\ 
\begin{table}[ht]
\caption{Measured Computer Live Times with Pulse Generator}
{\def\arraystretch{1.6}\tabcolsep=5pt
\begin{tabular}{c c c c}
\hline\hline
Pulse Rate & HMS & SHMS & BUSY width \\ [0.5ex] % inserts table %heading
\hline
100 Hz & 98.85$\%$ & 99.84 $\%$ &HMS: 2.2 ms \\
500 Hz & 50.07$\%$ & 98.30 $\%$  &SHMS: 2.2 $\mu$s \\
1 kHz  & 33.44$\%$ & 99.21 $\%$  \\
5 kHz  & 8.33$\%$ & 45.2 $\%$ \\
10 kHz & 4.30$\%$ & 30.49 $\%$ \\ [1ex]
\hline
\end{tabular}
}
\label{table:cp_LT}
\end{table} 
\indent From Eqs. \ref{eq:period} and \ref{eq:L.T.}, and the BUSY widths in Table \ref{table:cp_LT}, the expected computerlive times for certain frequencies can be calculated and the results are shown in
Tables \ref{table:expected_hmscp_LT} and \ref{table:expected_shmscp_LT}. \\
\begin{table}[ht]
\caption{Expected HMS Computer Live Times with Pulse Generator}
{\def\arraystretch{1.6}\tabcolsep=5pt
\begin{tabular}{c c c}
\hline\hline
($T=\frac{1}{n}w$) & Pulse Rate ($f = \frac{1}{T}$) & HMS ($\frac{1}{1+n}$) \\ [0.5ex] % inserts table %heading
\hline
$>w$             & $<$ 500 Hz & 100 $\%$ \\
$w$              & 500 Hz & 50  $\%$ \\
$\frac{1}{2}w$   & 1 kHz & 33.3 $\%$  \\
$\frac{1}{4}w$   & 2 kHz & 20.0 $\%$  \\
$\frac{1}{10}w$  & 5 kHz & 9.09 $\%$ \\
$\frac{1}{20}w$  & 10 kHz & 4.76 $\%$ \\ [1ex]
\hline
\end{tabular}
}
\label{table:expected_hmscp_LT}
\end{table} 
\begin{table}[ht]
\caption{Expected SHMS Computer Live Times with Pulse Generator}
{\def\arraystretch{1.6}\tabcolsep=5pt
\begin{tabular}{c c c}
\hline\hline
($T=\frac{1}{n}w$) & Pulse Rate ($f = \frac{1}{T}$) &  SHMS ($\frac{1}{1+n}$) \\ [0.5ex] % inserts table %heading
\hline
$>w$             & $<$ 454 kHz & 100 $\%$ \\
$w$              & 454 kHz & 50  $\%$ \\
$\frac{1}{2}w$   & 909 kHz & 33.3 $\%$  \\
$\frac{1}{4}w$   & 1.81 MHz & 20.0 $\%$  \\ [1ex]
\hline
\end{tabular}
}
\label{table:expected_shmscp_LT}
\end{table} 
\indent A comparison between the expected and measured computer live times in the HMS shows a discrepancy of $<$1$\%$ for rates $>500$Hz.
In the SHMS, however, since the BUSY signal is 10$^{3}$ orders of magnitude smaller than the HMS, a much high rate of 454 kHz is
expected before a dropping below 100$\%$ computer live times. Since the maximum rate measured in the SHMS was only 10 kHz, it was
expected that computer live times should have stayed at 100$\%$, but this was not the case. \\
\indent In addition to these measurements, to simulate physics events, a poisson pulse generator was also used to observe the
effects on the computer live times and view the signals on the Oscilloscope for instructional purposes (See Figure \ref{fig:scope_edtm}), but extensive measurements were
not done since the discrepancy between expected and measured livetimes using a pulse generator was a significant problem that needed to be
resolved. 
  \includegraphics[width=3.4in, height=2.6in]{edtm/EDTM_Oscope.pdf}
  \caption{Random (Poisson) and periodic (EDTM - elecreonic dead time monitoring) pulses initiating a TI BUSY as viewed in an
    Oscilloscope. The pulses appear as a line because the horizontal scale is set to 100 $\mu$s / division, and the pulses have
  widths on the order of tens of nanoseconds, so they appear $10^{3}$ smaller.}
  \label{fig:scope_edtm}
\end{figure}\\
Figure \ref{fig:scope_edtm} is a visual representation of how the TI module responds to various input signals. The yellow signals were produced by
a random pulse generator to replicate physics events, and the blue pulse is periodic. The first pulse (yellow) initiates a BUSY (pink) signal from the
TI, but no other signals arrived during that time. A second pulse (blue) initiates a second BUSY signals, and during the 220 $\mu$s that the window
is open, two other random pulses arrived, but are rejected by the TI BUSY, since the other pulse was still being processed.

\subsection{Technical Issues with the DAQ and Trigger Interface Module}
\noindent The discrepancies found in the computer live times using periodic pulse generator shed light on unknown problems regarding the DAQ and TI
module that required the intervention of experts. One of the problems had to do with SYNCHRONIZATION issues between the TI and external scaler modules
which are used to count triggers. When taking experimental data, one must go over a series of steps in order to initiate and end the run. These steps
involve the communication between the TI and Scaler modules during a START/END of run processes (See Figure \ref{fig:TI_sync}). 
  \includegraphics[width=3.4in, height=1.5in]{edtm/TI_Scaler_SYNC.pdf}
  \caption{Diagram showing unsynchronized TI and scaler modules.}
  \label{fig:TI_sync}
\end{figure}\\
When a run is started, a signal is sent to the TI module to initiate data processing, but the TI must also communicate with the scalers so that both
modules start counting simultaneously (in SYNC). The problem was that the TI and Scalers were not synchronized, so when a run was started, the scalers
would start counting events before the TI which caused a discrepancy in the live time calculations when a pulse generator was used. A second problem
was that the TI module internal scalers kept counting triggers after the END of run, which lead to an incorrect calculation of the computer live time
reported by the TI. \\
\indent These issues have been resolved, and the computer live time tests will continue on the near future to confirm that there are no other problems
with the live time calculations.
\section{D(e,e'p)n Experimental Run Plan}
\noindent My thesis subject, \textit{Deuteron Electro-Disintegration at Very High Missing Momenta} experiment is projected to run from February 26 to
March 3, 2017. The experiment consists of breaking the deuteron apart into a proton and neutron with an electron beam. The scattered electron and proton
will be detected by the SHMS and HMS, respectively, and the neutron will be reconstructed from mass-energy conservation laws. There will be a total of
five kinematics settings to be explored, two of which will be for spectrometer check and calibration purposes, and the other three settings will be the main
experimental settings that will explore the deuteron cross-section in unexplored regions of missing momentum with reduced final state interactions between
the proton and neutron. \\
\subsection{Spectrometer Settings: H(e,e'p) Scattering}
\noindent The first spectrometer setting will be to check that the spectrometers are pointing in the correct orientation in terms of angle and central
momentum. This is done via the elastic scattering of electrons off a 10 cm liquid Hydrogen target. In elastic scattering, e+p$\rightarrow$ e'+p, the
proton remains unchanged and re-scatters at a certain angle and momentum. By conservation laws, the final electron energy, $E'$, is determined to be
\begin{equation}\label{eq:h_elastic}
E'(\theta_{e}) = \frac{E_{e}}{1 + \frac{2E_{e}}{M_{p}}\sin^{2}\frac{\theta_{e}}{2}}
where $E_{e}$ is the initial electron energy, $M_{p}$ is the proton mass and $\theta_{e}$ is the electron scattering angle. From Eq. \ref{eq:h_elastic},
once the electron angle, $\theta_{e}$ is determined, the electron scattering energy as well as the proton scattering angle and momentum, $\theta_{p}$ and $P_{p}$
are also fixed. The D(e,e'p)n kinematics at high missing momentum require the electron angle to be fixed at $\theta_{e}$= 12.189$^{o}$, as it is the most
sensitive kinematic variable to the cross sections we want to measure. This way, one does not run the risk of changing the SHMS spectrometer angle during the
experimental run. The kinematics are summarized in Table \ref{Table:h_elastic_kin}.
\begin{table}[ht]
\caption{Hydrogen Elastic Kinematics}
{\def\arraystretch{1.6}\tabcolsep=5pt
\begin{tabular}{c c c}
\hline\hline
\textbf{Spectrometer Settings} & \textbf{HMS} & \textbf{SHMS} \\ [0.5ex] % inserts table %heading
\hline
particles                  & protons & electrons \\
Central Angle (deg)        & 37.339 & 12.169  \\
Central Momentum (GeV/c)   & 2.938 & 8.700  \\ [1ex]
\hline
\end{tabular}
}
\label{Table:h_elastic_kin}
\end{table} \\
\subsection{Spectrometer Settings: D(e,e'p)n Kinematics}
The spectrometers will be set to measure D(e,e'p)n kinematics at low missing momentum setting first,
since there exist similar data from which we can compare results and apply any
corrections needed before exploring the high missing momentum region. Three high missing momentum values will then be explored,
each corresponding to a different HMS spectrometer setting. During the entire D(e,e'p)n runs, the SHMS spectrometer must stay at a
fixed angle and central momentumn. The SHMS settings for detecting electrons are:
\begin{table}[ht]
\caption{SHMS Central Kinematic Settings}
{\def\arraystretch{1.6}\tabcolsep=5pt
\begin{tabular}{c c}
\hline\hline
\textbf{Spectrometer Settings} & \textbf{SHMS} \\ [0.5ex] % inserts table %heading
\hline
particles detected                  & electrons \\
Central Angle (deg)        & 12.169  \\
Central Momentum (GeV/c)   & 8.700  \\ [1ex]
\hline
\end{tabular}
}
\label{Table:shms_d2_kin}
\end{table} \\
The HMS settings for detecting protons at low and high missing momentum of neutron are:
\begin{table}[ht]
\caption{HMS Central Kinematic Settings}
{\def\arraystretch{1.6}\tabcolsep=2pt
\begin{tabular}{c c c c c}
\hline\hline
\textbf{Spectrometer Settings} & \textbf{HMS} \\ [0.5ex] % inserts table %heading
\hline
particles  detected                &           &    protons &  \\
\hline
$P_{miss}$ (GeV/c)           & 0.08     &  0.500  & 0.650 & 0.800\\
$\theta_{HMS}$ (deg)       & 40.061   & 53.252    & 56.401 & 59.388  \\    
$P_{HMS}$ (GeV/c)            & 2.435    & 2.305    & 2.220 & 2.120  \\
\hline
\end{tabular}
}
\label{Table:hms_d2_kin}
\end{table} \\
\subsection{Reaction Kinematics}
\indent The true electron reaction kinematics differ from the spectrometer settings described in Tables
\ref{Table:h_elastic_kin} and \ref{Table:shms_d2_kin}. The kinematics for the scattered electron are
\begin{table}[ht]
\caption{Electron Kinematic Settings}
{\def\arraystretch{1.6}\tabcolsep=5pt
\begin{tabular}{c c c}
\hline\hline
& \textbf{H(e,e'p)} & \textbf{D(e,e'p)n} \\ [0.5ex] % inserts table %heading
\hline
$\theta_{e}$ (deg)    & 12.169 & 12.169 \\           
$P^{'}_{e}$  (GeV/c)   & 8.454 & 8.922 \\
\hline
\end{tabular}
}
\label{Table:elec_kin}
\end{table} \\
In going from Hydrogen to Deuteron kinematics in the experimental run plan, the electron
momentum actually increases as shown in table \ref{Table:elec_kin}. We could have chosen the
SHMS central momentum settings to match the reaction kinematics, however, changing the central
momentum settings from low to high momentum in the quadrupole magnets leads to a phenomenon
